---
title: "01a.ordi.plots.MICEtraits"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(cowplot)
library(reshape2)
library(ggplot2)
library(ggrepel)
#library(grid)
#library(FD)
library(geiger)
library(data.table)
library(tidyr)
library(dplyr)
library(vegan)
library(ggforce)
```

#Ordination using P/A and sub traits (WITHOUT weighted average)
```{r}
head(df.PA)
traits.mat <- read.csv('../DATA/ImputedMiceTraits.2016-01-06.csv', as.is = T, row.names = 1)
setDT(traits.mat, keep.rownames = "Species")[]
traits.mat <- traits.mat %>% filter (Species %in% colnames(df.PA[8:ncol(df.PA)]))

#df.PA <- df.PA %>% select (-c(new.rep, Code))

df.PA.long <- melt(df.PA, id.vars = c("year", "plot", "type", "phyD", "fd"), variable.name = "Species")
df.PA.long <- df.PA.long[order(df.PA.long$plot),] #order by plot not species

plots.traits.mat <- df.PA.long %>% 
            left_join(traits.mat, by = "Species")

plots.traits.mat$photosyntheticPathway <- as.numeric(sub("C3", "1", plots.traits.mat$photosyntheticPathway))
plots.traits.mat$photosyntheticPathway[is.na(plots.traits.mat$photosyntheticPathway)] <- 0 #replace NA for C4 with 0
drop.traits <- c("floweringStartcoded", "lifeCycleCoded", "wis_IL", "Lifeform", "habitClean")
plots.traits.mat <- plots.traits.mat %>% select (-c(drop.traits))
head(plots.traits.mat)
plots.traits.mat.f <- filter(plots.traits.mat, value != "0")

#calcuate mean trait values across species per plot
mean.traitsByplot <- plots.traits.mat.f %>% 
group_by(plot, year, type, phyD, fd) %>%
  summarise_if(is.numeric, mean, na.rm = TRUE)

pa.tmds <- metaMDS(mean.traitsByplot[,-c(1:6)], k = 3, maxit = 25000, distance = "bray")
pa.tmds <- metaMDS(mean.traitsByplot[,-c(1:6)], k = 3, maxit = 25000, previous.best = pa.tmds)  
#stressplot(pa.tmds)
pa.scores <- as.data.frame(scores(pa.tmds)) #extract nmds scores
pa.scores2 <- cbind(pa.scores, as.data.frame(mean.traitsByplot[,1:5] ))

#pa.scores2$phyD <- data.frame(phyD = substr(pa.scores2$Code,1,1)) #split code
#pa.scores2$tD <- data.frame(tD =substr(pa.scores2$Code,2,2))

#split groupings for plotting
pa.scores3 <- pa.scores2[,-4]  #take out plot or below mean wont work
#pa.scores3$Plug <- subset(pa.scores3$Plug, select=-c(Code))
factor(pa.scores3$phyD, levels = c('L', 'M', 'H')) -> pa.scores3$phyD
factor(pa.scores3$fd, levels = c('L', 'H')) -> pa.scores3$fd

```

#Centroids
```{r}
centroids.pa.phyd <- subset(pa.scores3, select=-fd)%>% 
    group_by(year, type, phyD) %>% 
    summarise_all(funs(mean))

centroids.pa.fd<- subset(pa.scores3, select=-phyD) %>% 
    group_by(year, type, fd) %>% 
    summarise_all(funs(mean))
```


#TRAIT VECTORS
```{r}
#add trait vectors:
vf <- envfit(pa.tmds, mean.traitsByplot[,-c(1:6)], perm = 9999)
#vec <- as.data.frame(vf.Wtrait$vectors$arrows*sqrt(vf.Wtrait$vectors$r))
scores(vf, "vectors")
spp.scrs <- as.data.frame(scores(vf, display = "vectors"))
spp.scrs <- cbind(spp.scrs, Traits = rownames(spp.scrs))

```

#Plotting by fd
```{r}

centroids.fd <- subset(centroids.pa.fd, centroids.pa.fd$year == "2016" | centroids.pa.fd$year == "2019.09" )

start <- subset(pa.scores2, year == "2016")
end <- subset(pa.scores2, year == "2019.09")
dat2 <- rbind(start, end)

dat <-  subset(dat2, select=-type) %>%
   arrange(plot) %>%
  arrange(year)

dat %>%  
  ggplot() +
  geom_point(aes(x=NMDS1, y=NMDS2, shape=year, color = fd), size=3) +
  theme_bw() +
  labs(shape="Year", color="FD") +
  geom_path(aes(x = NMDS1, y = NMDS2, group = plot), arrow = arrow(length = unit(0.15, "cm")), color="grey") +
 #geom_segment(data = spp.scrs,             aes(x = 0, xend = NMDS1/3, y = 0, yend = NMDS2/3),     arrow = arrow(length = unit(0.25, "cm")), colour = "grey") +
  #geom_text_repel(data = spp.scrs, aes(x = NMDS1/3, y = NMDS2/3, label = Traits),  size = 3) +
  scale_color_manual("FD", values = c("L" = "#332288",
                                "H" = "#44AA99"
                                )) +
#stat_ellipse
     geom_mark_ellipse(expand = 0, data=start, aes(x=NMDS1,y=NMDS2, color= fd)) +
   geom_mark_ellipse(expand = 0, data = end, aes(NMDS1, NMDS2, color = fd)) + geom_path(data=centroids.fd, aes(x = NMDS1, y = NMDS2, group = fd), arrow = arrow(length = unit(0.15, "cm")), colour="red", size = 1)  

```


#Testing differences in distance traveled (vectors) by treatment
```{r}
pa.scores3
vector1 <- subset(pa.scores3, pa.scores3$year == "2016")
vector2 <- subset(pa.scores3, pa.scores3$year == "2019.09")

df <- cbind(vector1[,1:3], vector2[,1:3])
df$distance <- apply(df, 1, function(x) dist(matrix(x, nrow = 2, byrow = TRUE)))
cbind(df, vector1$fd) -> finalV
low <- subset(finalV, finalV$`vector1$fd` == "L")
high <- subset(finalV, finalV$`vector1$fd` == "H")
t.test(low$distance, high$distance)


#by PD
cbind(df, vector1$phyD) -> finalV.pd

lowPD <- subset(finalV.pd, finalV.pd$`vector1$phyD` == "L")
medPD <- subset(finalV.pd, finalV.pd$`vector1$phyD` == "M")
highPD <- subset(finalV.pd, finalV.pd$`vector1$phyD` == "H")
t.test(lowPD$distance, highPD$distance)
t.test(medPD$distance, highPD$distance)
t.test(medPD$distance, lowPD$distance)


```

#Dissimilarity by FD Treatment
```{r}

temp = lapply(split(mean.traitsByplot, mean.traitsByplot$year), function(x) split(x, x$fd))

dis.fd <- lapply(temp, function(x) lapply(x, function(x) vegdist(subset(x, select = seedMass:genome_Use), method="bray"))) 

n.dis.fd <- lapply(dis.fd, function(x) unlist(lapply(x, function(x) mean(x)) ) )

as.data.frame(n.dis.fd) -> diss.FD

names(diss.FD) <- c("2016", "2017", "2018.06", "2018", "2019.06", "2019")
rownames(diss.FD) -> diss.FD$fd
factor(diss.FD$fd, levels = c('L', 'H')) -> diss.FD$fd
melt(diss.FD) -> diss.FD.long
#test.fall <- subset(test, select = -c(`2018.05`, `2019.06`))
#test.spring <- subset(test, select = c(`2018.05`, `2019.06`))
#melt(test.fall) -> test.fall


#mean diss by FD and year plotted:
M <- diss.FD.long %>%
  mutate(group2 = case_when(variable %in% c("2018.06", "2019.06") ~ "A",
                            TRUE ~ "B")) %>% 
  ggplot(aes(x = variable, 
             y = value, 
             color = fd, 
             group = interaction(fd, group2), 
             linetype = group2)) +
  geom_line(size = 1) +
  scale_linetype_manual("season",
                        values = c("A" = "dashed",
                                   "B" = "solid"),
                        labels = c("A" = "spring",
                                   "B" = "fall")) + 
  scale_color_manual("FD",
                     values = c("L" = "#332288",
                                "H" = "#44AA99"
                                )) +
  geom_point(size = 3, alpha = 0.5) # only for comprehension, remove it
MM <- M + labs(x = "year",  y= "dissimilarity (Bray)", color="PD") + theme_bw() +
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))
  

```

#Confidence Intervals for dissimilarity
bootstrap original plots without replacement, recalculate diss
```{r}
##Matrices by year and phyD class:
temp = lapply(split(mean.traitsByplot, mean.traitsByplot$year), function(x) split(x, x$fd) )
#temp from above bootstrap 
##
myReps = 1000
myCI = 0.95
myError <- myObs <- vector('list', length = length(temp))
names(myError) <- names(myObs) <- names(temp)
for(i in names(myError)) myError[[i]] <- list(L = numeric(0), H = numeric(0))

for(i in names(temp)) {
  for(j in c('L', 'H')) {
    print(paste('doing year', i, 'with fd', j))
    bittyTemp <- temp[[i]][[j]][which(names(temp[[i]][[j]]) == 'seedMass'):which(names(temp[[i]][[j]]) == 'genome_Use')]
    bittyTemp.d <- sapply(seq(myReps), function(x) {
      vegdist(bittyTemp[sample(1:24, 12, replace = F), ], method = 'bray') %>% mean
    })
    myObs[[i]][[j]] <- mean(vegdist(bittyTemp, method = 'bray'))
    myError[[i]][[j]] <- c(obs = myObs[[i]][[j]], quantile(bittyTemp.d, c((1-myCI) / 2, myCI+((1-myCI) / 2)))) #half jacknifing in
    }
}

##################unnested lists and retaining the wide structure
myError -> err
#lapply(err$`2016`, function(x) as.data.frame(melt(x))) -> B
as.data.frame(melt(err)) -> ERR #need to keep the obs, 2.5, 
valueIDs <- c("obs","lowCI", "highCI" )
cbind(ERR, valueIDs) -> NEWdf.test

names(NEWdf.test) <- c("value", "fd", "year", "valueID")

data2fd <- dcast(NEWdf.test, year + fd ~ valueID, value.var = c("value"))
factor(data2fd$fd, levels = c('L', 'H')) -> data2fd$fd

MM <- data2fd %>%
  mutate(group2 = case_when(year %in% c("2018.05", "2019.06") ~ "A",
                            TRUE ~ "B")) %>% 
  ggplot(aes(x = year, 
             y = obs, 
             color = fd, 
             group = interaction(fd, group2), 
             linetype = group2)) +
  geom_line(size = 1) +
  scale_linetype_manual("season",
                        values = c("A" = "dashed",
                                   "B" = "solid"),
                        labels = c("A" = "spring",
                                   "B" = "fall")) + 
  scale_color_manual("FD",
                     values = c("L" = "#332288",
                                "H" =  "#44AA99"
                                )) +
  geom_point(size = 3, alpha = 0.5)  +
  geom_errorbar(aes(ymin=lowCI, ymax=highCI), width=.2,
                 position=position_dodge(.1)) 
  
MM <- MM + labs(x = "year",  y= "dissimilarity (Bray)", color="FD") + theme_bw() +
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))

fitA <- aov(obs ~ year + fd, data2fd)
summary(fitA)
TukeyHSD(fitA)


```










#Plot by PhyD
```{r}
dat %>%  
  ggplot() +
  geom_point(aes(x=NMDS1, y=NMDS2, shape=year, color = phyD), size=3) +
  theme_bw() +
  labs(shape="Year", color="PD") +
  geom_path(aes(x = NMDS1, y = NMDS2, group = plot), arrow = arrow(length = unit(0.15, "cm")), color="grey") +
  scale_color_manual("PD", values = c("L" = "#332288",
                                      "M" = "#117733",
                                "H" = "#44AA99"
                                )) +
  geom_mark_ellipse(data=start, aes(x=NMDS1,y=NMDS2, group= phyD, color= phyD), linetype="dashed") +
   geom_mark_ellipse(data = end, aes(NMDS1, NMDS2, group= phyD, color = phyD)) 

```


#Dissimilarity by PhyD Treatment
```{r}
temp = lapply(split(mean.traitsByplot, mean.traitsByplot$year), function(x) split(x, x$phyD))

dis.phyd <- lapply(temp, function(x) lapply(x, function(x) vegdist(subset(x, select = seedMass:genome_Use), method="bray"))) 

n.dis.phyd <- lapply(dis.phyd, function(x) unlist(lapply(x, function(x) mean(x)) ) )

as.data.frame(n.dis.phyd) -> diss.PD

names(diss.PD) <- c("planted", "year 1", "year 2.1", "year 2", "year 3.1", "year 3")
rownames(diss.PD) -> diss.PD$phyD
factor(diss.PD$phyD, levels = c('L','M', 'H')) -> diss.PD$phyD
melt(diss.PD) -> diss.PD.long
#test.fall <- subset(test, select = -c(`2018.05`, `2019.06`))
#test.spring <- subset(test, select = c(`2018.05`, `2019.06`))
#melt(test.fall) -> test.fall


#mean diss by PD and year plotted:
M <- diss.PD.long %>%
  mutate(group2 = case_when(variable %in% c("year 2.1", "year 3.1") ~ "A",
                            TRUE ~ "B")) %>% 
  ggplot(aes(x = variable, 
             y = value, 
             color = phyD, 
             group = interaction(phyD, group2), 
             linetype = group2)) +
  geom_line(size = 1) +
  scale_linetype_manual("season",
                        values = c("A" = "dashed",
                                   "B" = "solid"),
                        labels = c("A" = "spring",
                                   "B" = "fall")) + 
  scale_color_manual("FD",
                     values = c("L" = "#332288",
                                "M" = "#117733",
                                "H" = "#44AA99"
                                )) +
  geom_point(size = 3, alpha = 0.5) # only for comprehension, remove it
MM <- M + labs(x = "year",  y= "dissimilarity (Bray)", color="PD") + theme_bw() +
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))
  

```

#Confidence Intervals for dissimilarity
bootstrap original plots without replacement, recalculate diss
```{r}
##Matrices by year and phyD class:
temp = lapply(split(mean.traitsByplot, mean.traitsByplot$year), function(x) split(x, x$phyD) )
#temp from above bootstrap 
##
myReps = 1000
myCI = 0.95
myError <- myObs <- vector('list', length = length(temp))
names(myError) <- names(myObs) <- names(temp)
for(i in names(myError)) myError[[i]] <- list(L = numeric(0), H = numeric(0))

for(i in names(temp)) {
  for(j in c('L', "M", 'H')) {
    print(paste('doing year', i, 'with phyD', j))
    bittyTemp <- temp[[i]][[j]][which(names(temp[[i]][[j]]) == 'seedMass'):which(names(temp[[i]][[j]]) == 'genome_Use')]
    bittyTemp.d <- sapply(seq(myReps), function(x) {
      vegdist(bittyTemp[sample(1:24, 12, replace = F), ], method = 'bray') %>% mean
    })
    myObs[[i]][[j]] <- mean(vegdist(bittyTemp, method = 'bray'))
    myError[[i]][[j]] <- c(obs = myObs[[i]][[j]], quantile(bittyTemp.d, c((1-myCI) / 2, myCI+((1-myCI) / 2)))) #half jacknifing in
    }
}

##################unnested lists and retaining the wide structure
myError -> err
#lapply(err$`2016`, function(x) as.data.frame(melt(x))) -> B
as.data.frame(melt(err)) -> ERR #need to keep the obs, 2.5, 
valueIDs <- c("obs","lowCI", "highCI" )
cbind(ERR, valueIDs) -> NEWdf.test

names(NEWdf.test) <- c("value", "phyD", "year", "valueID")

data2phyd <- dcast(NEWdf.test, year + phyD ~ valueID, value.var = c("value"))
factor(data2phyd$phyD, levels = c('L',"M"  ,'H')) -> data2phyd$phyD

MM <- data2phyd %>%
  mutate(group2 = case_when(year %in% c("2018.05", "2019.06") ~ "A",
                            TRUE ~ "B")) %>% 
  ggplot(aes(x = year, 
             y = obs, 
             color = phyD, 
             group = interaction(phyD, group2), 
             linetype = group2)) +
  geom_line(size = 1) +
  scale_linetype_manual("season",
                        values = c("A" = "dashed",
                                   "B" = "solid"),
                        labels = c("A" = "spring",
                                   "B" = "fall")) + 
  scale_color_manual("FD",
                     values = c("L" = "#332288",
                                "M" = "#117733",
                                "H" =  "#44AA99"
                                )) +
  geom_point(size = 3, alpha = 0.5)  +
  geom_errorbar(aes(ymin=lowCI, ymax=highCI), width=.2,
                 position=position_dodge(.1)) 
  
MM <- MM + labs(x = "year",  y= "dissimilarity (Bray)", color="PD") + theme_bw() +
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))

fitA <- aov(obs ~ year + phyD, data2phyd)
summary(fitA)
TukeyHSD(fitA)


```
