---
title: "01b.ordi.plot.compositional.space"
output: html_document
---
#Shifts in Compositional Space
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(FD)
library(vegan)
library(ggplot2)
library(ggforce)
library(reshape2)
```

#Ordination of Plots by Composition
##Presence/Absence (Cover below)
```{r load data}
#source 00.load.for.analyses.Rmd: 
#DF1 <- as.matrix(df.pa.plug[8:length(df.pa.plug)])
DF1 <-as.matrix((subset(df.PA, select = Agalinis_tenuifolia:Zizia_aurea)))

jmds <- metaMDS(DF1, k=3, maxit=20000, distance = 'jaccard', autotransform = FALSE, trymax=100)
jmds <- metaMDS(DF1, distance = "jaccard", k=3, maxit=60000, previous.best=jmds, autotransform = FALSE, trymax=100) 
stressplot(jmds)
#scree.jmds <- dimcheckMDS(df.pa.plug[5:length(df.pa.plug)], distance = "jaccard", k = 6, trymax = 20,
#  autotransform = TRUE)


year = as.factor(df.PA$year)
plot = df.PA$plot
NMDS.plug.PAcomp <- as.data.frame(scores(jmds)) #NMDS.plug.PAcomp
NMDS.plug.PAcomp <- cbind(NMDS.plug.PAcomp, year, plot, df.PA$Code)
NMDS.plug.PAcomp$phyD <-  data.frame(phyD = substr(df.PA$Code,1,1)) 
unlist(NMDS.plug.PAcomp$phyD) -> NMDS.plug.PAcomp$phyD
factor(NMDS.plug.PAcomp$phyD, levels = c('L', 'M', 'H')) -> NMDS.plug.PAcomp$phyD
```

#Plot by PD treatment
```{r}
start <- subset(NMDS.plug.PAcomp, (year == "2016"))
end <- subset(NMDS.plug.PAcomp, (year == "2019.09"))
StoE <- rbind(start, end)

Fig2 <-  StoE %>%             #sort ascending so that 2019 is plotted 
  ggplot() +
  geom_point(aes(x=NMDS2, y=NMDS3, shape=year, color = phyD), size=3) +
  theme_bw() +
  labs(shape="Year", color="PD") +
  geom_path(aes(x = NMDS2, y = NMDS3, group = plot), arrow = arrow(length = unit(0.15, "cm")), color="grey") +
  
  stat_ellipse(data=start, aes(x=NMDS2,y=NMDS3, group = phyD), color= "grey") + 
  stat_ellipse(data=end, aes(x=NMDS2,y=NMDS3, color= phyD)) +
  scale_color_manual("PD", values = c("L" = "#332288",
                                "M" = "#117733",
                                "H" = "#44AA99"
                                )) +
annotate("text", x = 2.5, y = 2.7, color="grey",
label = "stress=0.22") + labs(x="nmds1", y="nmds2")

   
____________________
#SAME WITH FD
#NMDS.plug.PAcomp <- cbind(NMDS.plug.PAcomp, year, plot, df.PA$Code)

NMDS.plug.PAcomp$FD <-  data.frame(phyD = substr(df.PA$Code,2,2)) 
unlist(NMDS.plug.PAcomp$FD) -> NMDS.plug.PAcomp$FD
factor(NMDS.plug.PAcomp$phyD, levels = c('L', 'H')) -> NMDS.plug.PAcomp$FD
startFD <- subset(NMDS.plug.PAcomp, (year == "2016"))
endFD <- subset(NMDS.plug.PAcomp, (year == "2019.09"))
StoEFD <- rbind(startFD, endFD)

Fig2c <-  StoEFD %>%             #sort ascending so that 2019 is plotted 
  ggplot() +
  geom_point(aes(x=NMDS1, y=NMDS2, shape=year, color = FD), size=3) +
  theme_bw() +
  labs(shape="Year", color="FD") +
  geom_path(aes(x = NMDS1, y = NMDS2, group = plot), arrow = arrow(length = unit(0.15, "cm")), color="grey") +
  
  stat_ellipse(data=startFD, aes(x=NMDS1,y=NMDS2, group = FD), color= "grey") + 
  stat_ellipse(data=endFD, aes(x=NMDS1,y=NMDS2, color= FD)) +
  scale_color_manual("FD", values = c("L" = "#332288",
                                "M" = "#117733",
                                "H" = "#44AA99"
                                )) +
annotate("text", x = 2.5, y = 2.7, color="grey",
label = "stress=0.22") + labs(x="nmds1", y="nmds2")




####Ordination with abundance/cover data 
```{r}
nmds.plug.comp <- metaMDS(df.cover[8:ncol(df.cover)], k=3, maxit=20000, distance="bray", autotransform = FALSE, trymax=100)
nmds.plug.comp <- metaMDS(df.cover[8:length(df.cover)], k=3, maxit=20000, distance = 'bray', previous.best = nmds.plug.comp, autotransform = FALSE, trymax=100)

####Plot Plug plots Cover Ordination
year <- as.factor(df.cover.plug$year)
plot <- df.cover.plug$plot
NMDS.plug.comp <- as.data.frame(scores(nmds.plug.comp)) 
NMDS.plug.comp <- cbind(NMDS.plug.comp, year, plot, df.cover.plug$phyD)
names(NMDS.plug.comp) <- c("NMDS1", "NMDS2", "NMDS3", "year", "plot", "phyD")
factor(NMDS.plug.comp$phyD, levels = c('L', 'M', 'H')) -> NMDS.plug.comp$phyD
      ###NEED TO RENAME mds.plug to nmds.plug.comp #used for 03b.randomization
      #mds.plug -> mds.plug.comp

FigPlugCoverOrd <- NMDS.plug.comp %>% 
   arrange(plot) %>% #sort ascending so that 2018 is plotted last
  arrange(year) %>%
ggplot() +
  geom_point(aes(x=NMDS1, y=NMDS3, col=year, shape=phyD), size=3) +
  theme_bw() +
  labs(shape="PD", color="Year", x="nmds1", y="nmds2") +
  geom_path(aes(x = NMDS1, y = NMDS3, group = plot), arrow = arrow(length = unit(0.15, "cm")), color="grey") +
  stat_ellipse(aes(x=NMDS1, y=NMDS3, color = year, group = year), size=1)

FigPlugCoverOrd <- FigPlugCoverOrd + scale_color_manual(values=c('#999933','#44AA99', '#117733', '#CC6677', '#332288')) + annotate("text", x = 1.4, y = 1.3, color="grey", label = "stress=0.22")

#ggsave(FigPlugCoverOrd, file="../OUT/Supplement_CoverCompOrd.pdf", width=6, height=5, unit="in")

##Subet data for fall only>
subNMDS.cov <- subset(NMDS.plug.comp, !(year == "2018.05"))
subNMDS.cov <- subset(subNMDS.cov, !(year=="2019.06"))

FigCoverFall <- subNMDS.cov %>% 
   arrange(plot) %>% #sort ascending so that 2018 is plotted last
  arrange(year) %>%
ggplot() +
  geom_point(aes(x=NMDS1, y=NMDS3, col=year, shape=phyD), size=3) +
  theme_bw() +
  labs(shape="PD", color="Year", x="nmds1", y="nmds2") +
  geom_path(aes(x = NMDS1, y = NMDS3, group = plot), arrow = arrow(length = unit(0.15, "cm")), color="grey") +
  stat_ellipse(aes(x=NMDS1, y=NMDS3, color = year, group = year), size=1)

FigCoverFall <- FigCoverFall + scale_color_manual(values=c('#999933','#332288', '#117733')) + annotate("text", x = 1.4, y = 1.3, color="grey", label = "stress=0.22")
```
#COVER figure start to finish as above for supplement
startC <- subset(NMDS.plug.comp, (year == "2017"))
endC <- subset(NMDS.plug.comp, (year == "2019.09"))

StoECov <- rbind(startC, endC)

SupplementFig <-  StoECov %>%             #sort ascending so that 2019 is plotted 
  ggplot() +
  geom_point(aes(x=NMDS1, y=NMDS2, shape=year, color = phyD), size=3) +
  theme_bw() +
  labs(shape="Year", color="PD") +
  geom_path(aes(x = NMDS1, y = NMDS2, group = plot), arrow = arrow(length = unit(0.15, "cm")), color="grey") +
  
  stat_ellipse(data=startC, aes(x=NMDS1,y=NMDS2, group = phyD), color= "grey") + 
  stat_ellipse(data=endC, aes(x=NMDS1,y=NMDS2, color= phyD)) +
  scale_color_manual("PD", values = c("L" = "#332288",
                                "M" = "#117733",
                                "H" = "#44AA99"
                                )) +
                                labs(x="nmds1", y="nmds2")
#annotate("text", x = 2.5, y = 2.7, color="grey",label = "stress=0.22") + 


#Testing for Convergence/Divergence: Mean Compositional dissimilarity for P/A
####Plug and Seed
Comparing mean dissimilarity between years can indicate if plots are converging 
```{r}
#P/A Dissimilarity by year
df.pa.plug.yr <- split(df.pa.plug, df.pa.plug$year) 

dis.matJ.p <- lapply(df.pa.plug.yr, function(x) vegdist(x[8:ncol(x)], method="jaccard"))
mean.J.p <- unlist(lapply(dis.matJ.p, function(x) mean(x)))

#########################Dissimilarity by PhyD Class###################
temp = lapply(split(df.pa.plug, df.pa.plug$year), function(x) split(x, x$phyD))
nested.dis.matJ.p <- lapply(temp, function(x) lapply(x, function(x) vegdist(subset(x, select = Agalinis_tenuifolia:Zizia_aurea), method="jaccard"))) 

nested.mean.J.p <- lapply(nested.dis.matJ.p, function(x) unlist(lapply(x, function(x) mean(x)) ) )

as.data.frame(nested.mean.J.p) -> diss.plug
names(diss.plug) <- names(nested.mean.J.p)
rownames(diss.plug) -> diss.plug$phyd
factor(diss.plug$phyd, levels = c('L', 'M', 'H')) -> diss.plug$phyd
melt(diss.plug) -> diss.plug.long
#test.fall <- subset(test, select = -c(`2018.05`, `2019.06`))
#test.spring <- subset(test, select = c(`2018.05`, `2019.06`))
#melt(test.fall) -> test.fall

P <- diss.plug.long %>%
  mutate(group2 = case_when(variable %in% c("2018.05", "2019.06") ~ "A",
                            TRUE ~ "B")) %>% 
  ggplot(aes(x = variable, 
             y = value, 
             color = phyd, 
             group = interaction(phyd, group2), 
             linetype = group2)) +
  geom_line(size = 1) +
  scale_linetype_manual("season",
                        values = c("A" = "dashed",
                                   "B" = "solid"),
                        labels = c("A" = "spring",
                                   "B" = "fall")) + 
  scale_color_manual("PD",
                     values = c("L" = "#332288",
                                "M" = "#117733",
                                "H" = "#44AA99"
                                )) +
  geom_point(size = 3, alpha = 0.5) # only for comprehension, remove it
P <- P + labs(x = "year",  y= "dissimilarity (Jaccard's)", color="PD") + theme_bw() +
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))
  
  #labs(title = "Mean dissimilarity",subtitle = "Presence/absence data", x = "year",  y= "Dissimilarity (Jaccard's)") + theme_bw() +
#ggsave(P, file="../OUT/meanDissPA.png", width = 8, height = 6)
```

#Confidence Intervals for dissimilarity
bootstrap original plots without replacement, recalculate diss
```{r}
##Matrices by year and phyD class:
temp = lapply(split(df.pa.plug, df.pa.plug$year), function(x) split(x, x$phyD) )
#temp from above bootstrap 
##
myReps = 1000
myCI = 0.95
myError <- myObs <- vector('list', length = length(temp))
names(myError) <- names(myObs) <- names(temp)
for(i in names(myError)) myError[[i]] <- list(L = numeric(0), M = numeric(0), H = numeric(0))

for(i in names(temp)) {
  for(j in c('L', 'M', 'H')) {
    print(paste('doing year', i, 'with phyD', j))
    bittyTemp <- temp[[i]][[j]][which(names(temp[[i]][[j]]) == 'Agalinis_tenuifolia'):which(names(temp[[i]][[j]]) == 'Zizia_aurea')]
    bittyTemp.d <- sapply(seq(myReps), function(x) {
      vegdist(bittyTemp[sample(1:24, 12, replace = F), ], method = 'jaccard') %>% mean
    })
    myObs[[i]][[j]] <- mean(vegdist(bittyTemp, method = 'jaccard'))
    myError[[i]][[j]] <- c(obs = myObs[[i]][[j]], quantile(bittyTemp.d, c((1-myCI) / 2, myCI+((1-myCI) / 2)))) #half jacknifing in
    }
}

##################unnested lists and retaining the wide structure
myError -> err
#lapply(err$`2016`, function(x) as.data.frame(melt(x))) -> B
as.data.frame(melt(err)) -> ERR #need to keep the obs, 2.5, 
valueIDs <- c("obs","lowCI", "highCI" )
cbind(ERR, valueIDs) -> NEWdf.test

names(NEWdf.test) <- c("value", "phyd", "year", "valueID")

data2 <- dcast(NEWdf.test, year + phyd ~ valueID, value.var = c("value"))
factor(data2$phyd, levels = c('L', 'M', 'H')) -> data2$phyd

#myError has each obs and error bars
PP <- data2 %>%
  mutate(group2 = case_when(year %in% c("2018.05", "2019.06") ~ "A",
                            TRUE ~ "B")) %>% 
  ggplot(aes(x = year, 
             y = obs, 
             color = phyd, 
             group = interaction(phyd, group2), 
             linetype = group2)) +
  geom_line(size = 1) +
  scale_linetype_manual("season",
                        values = c("A" = "dashed",
                                   "B" = "solid"),
                        labels = c("A" = "spring",
                                   "B" = "fall")) + 
  scale_color_manual("PD",
                     values = c("L" = "#332288",
                                "M" =  "#117733",
                                "H" =  "#44AA99"
                                )) +
  geom_point(size = 3, alpha = 0.5)  +
  geom_errorbar(aes(ymin=lowCI, ymax=highCI), width=.2,
                 position=position_dodge(.1)) 
  
PP <- PP + labs(x = "year",  y= "dissimilarity (Jaccard's)", color="PD") + theme_bw() +
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))


fitA <- aov(obs ~ year + phyd, data2)
fitA <- aov(obs ~ year + phyd + year:phyd, data2)

summary(fitA)
TukeyHSD(fitA)


t.test()

```

#PLOT THE ORD AND CI together:
```{r}
cowplot::plot_grid(PP, FigPlugPACompOrd, labels = "auto", nrow = 1) -> Figure4.PAComposition.Plug
Figure4.PAComposition.Plug
#ggsave(file="../OUT/Figure2.PAComposition.pdf")
```

##########################Mean pairwise replicates
```{r}
#per $year, group_by phyD level, dist(), extract replicate dists only, calculate mean & SEM
df3 <- vector('list', length(df.pa.plug.yr))
names(df3) <- names(df.pa.plug.yr)
for(i in names(df.pa.plug.yr)) {
  df.pa.plug.yr[[i]] -> MM
  ## at the end, shove into df3[[i]]
  rownames(MM) <- paste(MM$plot, MM$new.rep, MM$phyD, sep="-")
  disMat <- vegdist(MM[8:ncol(MM)], method="jaccard")
  datmat <- as.matrix(disMat)
  library(reshape2)
  df3[[i]] <- melt(datmat)
  df3[[i]]$Var1 <- as.character(df3[[i]]$Var1)
  df3[[i]]$Var2 <- as.character(df3[[i]]$Var2)
  df3[[i]]$spMix1 <- sapply(strsplit(df3[[i]]$Var1, '-'), '[', 2)
  df3[[i]]$spMix2 <- sapply(strsplit(df3[[i]]$Var2, '-'), '[', 2)
  df3[[i]]$plot1 <- sapply(strsplit(df3[[i]]$Var1, '-'), '[', 1)
  df3[[i]]$plot2 <- sapply(strsplit(df3[[i]]$Var2, '-'), '[', 1)
  df3[[i]]$phyD <- sapply(strsplit(df3[[i]]$Var1, '-'), '[', 3)
  df3[[i]] <- df3[[i]][(df3[[i]]$plot1 != df3[[i]]$plot2) & (df3[[i]]$spMix1 == df3[[i]]$spMix2), ]
  
  df3[[i]] <- df3[[i]][!duplicated(df3[[i]]$spMix1), ] #drop duplicate pairwise

}

df3.subset <- df3.subset[!duplicated(df3.subset$spMix1), ]
df.pa.plug.yr$'2017'$phyD[match(df3.subset$plot2, df.pa.plug.yr$`2017`$plot)] #drop those matching

```


#Supplmental figures
#remove spring survey points
subNMDS.plug.PAcomp <- subset(NMDS.plug.PAcomp, !(year == "2018.05"))
subNMDS.plug.PAcomp <- subset(subNMDS.plug.PAcomp, !(year=="2019.06"))

FigPlugPACompOrd <- subNMDS.plug.PAcomp %>% 
arrange(plot) %>% 
arrange(year) %>%             #sort ascending so that 2018 is plotted last (for arrows)
  ggplot() +
  geom_point(aes(x=NMDS1, y=NMDS3, col=year,shape=phyD), size=3) +
  theme_bw() +
  labs(shape="PD", color="Year") +
  geom_path(aes(x = NMDS1, y = NMDS3, group = plot), arrow = arrow(length = unit(0.15, "cm")), color="grey") +
  stat_ellipse(aes(x=NMDS1, y=NMDS3, color = year, group = year), size=1)

FigPlugPACompOrd <- FigPlugPACompOrd + scale_color_manual(values=c('#88CCEE','#999933', '#117733', '#332288')) + annotate("text", x = 2.5, y = 2.7, color="grey", label = "stress=0.22") + labs(x="nmds1", y="nmds2")

#ggsave(FigPlugPACompOrd, file="../OUT/FigPlugPACompOrd_fall.jpeg", width=6, height=5, units="in")

####################Now 2017 to spring only
subNMDS.spring.PAcomp <- subset(NMDS.plug.PAcomp, !(year == "2018.09"))
subNMDS.spring.PAcomp <- subset(subNMDS.spring.PAcomp, !(year=="2019.09"))

FigPACompOrdSPRING <- subNMDS.spring.PAcomp %>% 
arrange(plot) %>% 
arrange(year) %>%             #sort ascending so that 2018 is plotted last (for arrows)
  ggplot() +
  geom_point(aes(x=NMDS1, y=NMDS3, col=year,shape=phyD), size=3) +
  theme_bw() +
  labs(shape="PD", color="Year") +
  geom_path(aes(x = NMDS1, y = NMDS3, group = plot), arrow = arrow(length = unit(0.15, "cm")), color="grey") +
  stat_ellipse(aes(x=NMDS1, y=NMDS3, color = year, group = year), size=1)

FigPACompOrdSPRING  <- FigPACompOrdSPRING + scale_color_manual(values=c('#88CCEE','#999933', '#44AA99', '#CC6677')) + annotate("text", x = 2.5, y = 2.7, color="grey", label = "stress=0.22") +
  labs(x="nmds1", y="nmds2")

#'#999933','#88CCEE', '#CC6677', '#117733', '#882255', '#332288'
#ggsave(FigPACompOrdSPRING, file="../OUT/Supplement_PACompOrd_spring.png", width=6, height=5, unit="in")

