---
title: "03.Randomization.traits"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vegan)
library(reshape2)
library(permute)
library(magrittr)
library(scatterplot3d)
library(tidyr)
```

#Randomization test of shifts/points in TRAIT space
Given the points from ordination trait space by weighted average from start to finish, use that vector of change per plot as the radius and randomly generate new points on the sphere. Test if the direction across plots (for convergence or similar shifts) is significant.

##DATA from ordinations
Note: Run with both P/A and cover data
```{r plug trait data}
#run 01a.ordi.plot.trait.space for non-MICE weighted.traits ordination
#01a.ordi.plot.MICEtraits.Rmd =  Wscores
#01a.ordi.plot.MICEtraits.Rmd P/A data 

#Wscores.split <- split(Wscores, Wscores$type)
Wscores.pl <- split(Wscores, Wscores$year)
start <- as.data.frame(Wscores.pl$`2017`[,5:7]) 
end <- as.data.frame(Wscores.pl$`2019.09`[,5:7])
```

#Sphere Function
```{r}
#Function to generate x,y,z points given start point (center)
sphere <- function (sim.s=NA, sim.t=NA, addTo = NA, r = 10) {
# addTo is the 2017 nMDS matrix (Ssplit17 or plug.scores17)
	if(is.na(sim.s)) {
		sim.s = runif(length(r), 0, 359.9999)
		sim.t = runif(length(r), 0, 359.9999)
		}
    sim.s <- sim.s * pi/180
    sim.t <- sim.t * pi/180
	out <- cbind(x = r * cos(sim.s) * sin(sim.t), y = r * sin(sim.s) * sin(sim.t), 
        z = r * cos(sim.t))
	if(!is.na(addTo[1])) out <- out + addTo
    return(out)
}
#function to give you the radius for the above
radii <- function(m1, m2, cols=1:3) {
  m3 <- m1[,cols] - m2[,cols]
  out <- apply(m3, 1, function(x) sqrt(sum(x^2)))
out 
}
```

#Simulate plot positions
```{r}
#run sphere and radii functions above then:
sims <- lapply(1:1000, function(x) {
            sphere(addTo = as.matrix(start[,1:3]), r=radii(start, end) )
}
)
#plot simulated points and observed start - end
lapply(sims, function(x) unlist(x[,1])) -> sims.x #all my simulated x points
unlist(sims.x) -> sims.x
lapply(sims, function(x) unlist(x[,2])) -> sims.y
unlist(sims.y) -> sims.y
lapply(sims, function(x) unlist(x[,3])) -> sims.z
unlist(sims.z) -> sims.z
simSXYZ <- cbind(sims.x, sims.y, sims.z)

a <- scatterplot3d(simSXYZ, xlim=c(-0.4,0.4), ylim=c(-0.4,0.4), zlim =c(-0.4,0.4), color = "lightgrey")
a$points3d(start[,1:3], col="black", pch=19)
a$points3d(end[,1:3], col="blue", pch=19)
a <- scatterplot3d(simSXYZ, xlim=c(-0.4,0.4), ylim=c(-0.4,0.4), zlim =c(-0.4,0.4), angle = 50)

a <- scatterplot3d(simSXYZ)
a$points3d(Wscores.pl$`2017`[,1:3], col="green")
a$points3d(Wscores.pl$`2019.09`[,1:3], col="blue")
```

Are start - finish plots closer in trait space (constrained) than expected given the simulated data?
To test for directional movement in trait space - test for significant differences in distances of centroid. 
NULL: plots move only the same distance but any direction. DIRECTIONALITY: Test distance of centroids to infer cloud movement. 

#DISTANCE OF CENTROIDS 
###Mean Obs start - obs end vs simulated end
Compare mean distance vectors of observed and simulated by plotting distribution of average movement to start centroid compare to that of end
Find center point of all plots in trait space
```{r}
#Calculate mean dissimilarity for observed data:
mean(dist(start[,1:3])) -> mean.start.dis  
mean(dist(end[,1:3])) -> mean.obs.end.dis 

#dist() functions row by row - we want mean(dist() for sim[[1]]:[[1000]]) one mean for each 1000 simulation
lapply(sims, dist) -> Dp #distance matrices
lapply(Dp, mean) -> Mp #mean Distances for simulated 
unlist(Mp)-> Mp
#plot distribution of mean simulated plot dissimilarities 
hist(Mp, main="Mean simulated dissimilarities") #, xlim= c(2.4, 2.9)) 
abline(v=mean.start.dis, lwd=2,col="green")
abline(v=mean.obs.end.dis, lwd=2,col="blue")
min(sum(mean.obs.end.dis <= Mp), (sum(mean.obs.end.dis >= Mp)) *2) / (length(Mp)+1) 
```


```{r centroid}
#Calculate mean points for start and end observed data:
mean.start.centroid <- apply(Wscores.pl$`2017`[,5:7], 2, mean) #modify as needed
mean.end.centroid <- apply(Wscores.pl$`2019.09`[,5:7], 2, mean)
dist(rbind(mean.start.centroid, mean.end.centroid)) -> obs.centroid.dist 

#center without x,y,z seperately:
a <- scatterplot3d(Wscores.pl$`2017`[,5:7])
matrix(mean.start.centroid, nrow=1) -> b
a$points3d(b, col="red", pch=19)
matrix(mean.end.centroid, nrow=1) -> c
a$points3d(c, col="blue", pch=19)

sim.centroids <- t(sapply(sims, function(x) apply(x, 2, mean)))
apply(sim.centroids, 1, function(x) list(rbind(x, mean.start.centroid))) -> test3
test3 <- lapply(test3, '[[', 1)
lapply(test3, function(x) dist(as.data.frame(x)) ) -> vectors
unlist(vectors) -> vectors
hist(vectors)
abline(v=obs.centroid.dist, col="green")

min(sum(obs.centroid.dist <= vectors), (sum(obs.centroid.dist >= vectors)) *2) / (length(vectors)+1)
```

#Plot simulated centroids
```{r}
as.data.frame(sim.centroids) -> sim.centroidsAt 
#Wscores.split <- split(Wscores, Wscores$type)
#Wscores.pl <- split(Wscores.split$Plug, Wscores.split$Plug$year)
#mean.start.centroid <- apply(Wscores.pl$`2017`[,6:8], 2, mean)
#mean.end.centroid <- apply(Wscores.pl$`2019.09`[,6:8], 2, mean)

SIMt <- ggplot()+
theme_bw() +
geom_point(data=sim.centroidsAt, aes(x=x, y=y), color="lightgrey") + #2016-2019
#stat_ellipse(data=sim.centroidsA, aes(x=x, y=y), color="darkgrey") 
geom_point(data=as.data.frame(t(mean.start.centroid)), aes(x= NMDS1, y= NMDS2), color="black", size=5) +
annotate("text", x = -0.01, y = -0.0085, label = "2016 Centroid", size=6) +
geom_point(data=as.data.frame(t(mean.end.centroid)), aes(x= NMDS1, y= NMDS2), color="black", size=3) +
annotate("text", x =-0.015, y = 0.0035, label= "2019.09 Centroid", size=6) +
  labs(y="nmds2", x="nmds1") +
 annotate("text", x =0.0022, y = 0.003, label= "p=0.000999", size=6, color="darkgrey") 
 # labs(title="Centroid positions", subtitle = "Simulated for 2019.09", y="nmds2", x="nmds1")
  
SIMt
#ggsave(file="../OUT/Traitsimulated.ordinations.2016-2019.plug.png")
```