---
title: "simper_traitspace_Phylo"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(cowplot)
library(vegan)
library(data.table)
library(ggtree)
library(ggplot2)
library(dplyr)
library(ape)
library(ggrepel)
library(dplyr)
library(reshape2)
```
#SIMPER:Contribution of individual species to the overall dissimilarity
```{r}
df.PA[, 1:7] -> df.env
sim <- with(df.env, simper(df.PA[,8:ncol(df.PA)], year, permutations = 1000))
sim
summary(sim) -> simper.out
contribution.value <- cbind(rownames(simper.out$`2016_2019.09`), simper.out$`2016_2019.09`$average)
as.data.frame(contribution.value) -> contribution.value
names(contribution.value) <- c("species", "cumsum")
contribution.value.ordered <- contribution.value[order(contribution.value$species),] 

#SIMPER species in trait space
traits.mat <- read.csv('../DATA/ImputedMiceTraits.2016-01-06.csv', row.names = 1)
setDT(traits.mat, keep.rownames = "Species")[]
#replace "Symphyotrichum_novae.angliae" to "Symphyotrichum_novae-angliae"
traits.mat$Species <- gsub("Symphyotrichum_novae-angliae", "Symphyotrichum_novae.angliae", traits.mat$Species)

traits.mat <- traits.mat %>% filter (Species %in% contribution.value.ordered$species) 
traits.mat$photosyntheticPathway <- as.numeric(sub("C3", "1", traits.mat$photosyntheticPathway))
traits.mat$photosyntheticPathway[is.na(traits.mat$photosyntheticPathway)] <- 0 #replace NA for C4 with 0
rownames(test.traits) <- test.traits$Species
drop.traits <- c("floweringStartcoded", "lifeCycleCoded", "wis_IL", "Lifeform", "habitClean", "Species")
new.traits.mat <- traits.mat %>% select (-c(drop.traits))

Ord <- metaMDS(new.traits.mat, maxit=10000, k=3, distance = "gower") #stress = 0.15

Species.ord <- as.data.frame(scores(Ord)) 
rownames(Species.ord) -> Species.ord$Species
#add in Percent Change in veg cover to color code 
#pct_changeEnd #from 02a.winners.losers.Rmd is mean cover change
cbind(Species.ord, contribution.value.ordered, as.data.frame(pct_changeEnd)) -> species.dat 
subset(species.dat, select=-c(Species)) -> species.dat
as.numeric(as.character(species.dat$cumsum)) -> species.dat$cumsum #species.dat is the nmds matrix.
#add a type column based on  species.dat, cumsum > 0.02 which indicates SIMPER
species.dat -> df
df <- df %>% mutate(NewCol = ifelse(cumsum >= 0.0123, "simper", "no"))

species.in.space <- ggplot(species.dat, aes(NMDS1, NMDS2, size=cumsum, color=PctChange, label=species)) +
   geom_point() + 
   scale_color_gradient2(low = "blue", mid = "white",
  high = "red", midpoint = 0) +
  labs(color = "cover change", size= "simper") + 
  theme_bw()
species.in.space2 <- species.in.space  +
  geom_text_repel(data=subset(species.dat, cumsum >= 0.0123), size=2.5, vjust=2, color="black", min.segment.length = unit(0.01, 'lines'),nudge_y = 0) 

#pc = data used in nmds: 
pc <- test.traits[-1]
pc2 <- left_join(test.traits, df2, by="species")
m_com = as.matrix(pc)

ano = anosim(m_com, pc2$simperType, distance = "gower", permutations = 9999)

d <- dist(test.traits[-1], method="euclidean") #dissimilarity matrix
clust <- hclust(d, method = "complete", members = NULL)
plot(Ord) #plot ordination points. red points = traits
ordicluster(Ord, clust, prune=3, col = cutree(clust, h=6)) #add vectors 4 colors
#use clust, idenfiy groups (based on X and see if simper are within the group)
#subset clust$labels for the simper only
#subset(species.dat, cumsum > 0.02) -> sim
clust$labels %in% sim$Species #returns logical use to subset $labels and $heights

clust[clust$labels %in% sim$Species, ]
subset(clust$labels, clust$labels %in% sim$Species)
#plot(Ord, type='t', display=c('species')) plots traits only

```

```{r}
phy <- read.tree("../DATA/phylogeny.analyzed.2016-01-05b.tre") 
phy$tip.label    #140
phy$tip.label[which(phy$tip.label == "Symphyotrichum_novae-angliae")] <-
  "Symphyotrichum_novae.angliae"
tree <- drop.tip(phy, which(!phy$tip.label %in% rownames(species.dat))) 

rownames(species.dat) <- species.dat$Species
setDT(species.dat, keep.rownames = "Species")[] -> species.dat
species.dat <- as.data.frame(species.dat)

p <- ggtree(tree, layout="circular")

###
species.dat <- species.dat %>%
  select(Species, everything())

pp2 <- p %<+% species.dat + geom_tippoint(aes(color=change, offset = 5)) + 
  scale_color_gradient2(low = "blue", mid = "white",
  high = "red", midpoint = 0) +
    theme(legend.position = "none") +
  labs(title = "", color = "Cover Change 2017-2019", size= "simper Contribution") 

p.cumsum <- p %<+% species.dat + geom_tippoint(aes(color=cumsum, offset = 5), size=3) + 
  scale_color_gradient2(low = "white",
  high = "black") +
    theme(legend.position = "none") +
  labs(title = "", color = "Cover Change 2017-2019", size= "simper Contribution") 

p2 <- p %<+% species.dat + geom_tippoint(aes(color=change, size=cumsum, offset = 2)) + 
  scale_color_gradient2(low = "blue", mid = "white",
  high = "red", midpoint = 0) +
    theme(legend.position = "none") + scale_size_continuous(range = c(0, 6) ) +
  labs(title = "", size = "simper contribution", color = "Cover Change 2017-2019", size= "simper Contribution") 

p3 <- p2 + geom_cladelabel(node=192, label="Asteraceae", offset = 5) +
            geom_cladelabel(node=233, label="Lamiaceae", offset = 5, hjust = 1.1) +
            geom_cladelabel(node=173, "Fabaceae", offset = 5, hjust = 1) +
            geom_cladelabel(node=169, "Rosaceae", offset = 5, hjust =1.1) +
            geom_cladelabel(node=147, "Poaceae", offset = 5, hjust = -.1)

p3  

plot_grid(species.in.space2, p3, rel_widths = c(1.5, 1),labels = "auto") -> simper.fig
simper.fig
```


#SIMPER SPECIES in trait space based on Vegetative Cover
```{r}
############COVER#####################
df.cover[,1:7] -> df.env #made in 00b.load.for.analyses
sim <- with(df.env, simper(df.cover[,8:ncol(df.cover)], year, permutations = 1000))
sim
summary(sim) -> simper.out

contribution.value <- cbind(rownames(simper.out$`2017_2019.09`), simper.out$`2017_2019.09`$average)
as.data.frame(contribution.value) -> contribution.value
names(contribution.value) <- c("species", "cumsum")
contribution.value.ordered <- contribution.value[order(contribution.value$species),] 
#
traits.mat <- read.csv('../DATA/ImputedMiceTraits.2016-01-06.csv', row.names = 1)
setDT(traits.mat, keep.rownames = "Species")[]
#replace "Symphyotrichum_novae.angliae" to "Symphyotrichum_novae-angliae"
traits.mat$Species <- gsub("Symphyotrichum_novae-angliae", "Symphyotrichum_novae.angliae", traits.mat$Species)
traits.mat <- traits.mat %>% filter (Species %in% contribution.value.ordered$species) 
traits.mat$photosyntheticPathway <- as.numeric(sub("C3", "1", traits.mat$photosyntheticPathway))
traits.mat$photosyntheticPathway[is.na(traits.mat$photosyntheticPathway)] <- 0 #replace NA for C4 with 0
rownames(test.traits) <- test.traits$Species

drop.traits <- c("floweringStartcoded", "lifeCycleCoded", "wis_IL", "Lifeform", "habitClean", "Species")
new.traits.mat <- traits.mat %>% select (-c(drop.traits))

Ord <- metaMDS(new.traits.mat, maxit=10000, k=3, distance= "gower") #stress = 0.135

Species.ord <- as.data.frame(scores(Ord)) 
rownames(Species.ord) -> Species.ord$Species
#Percent change in cover
pct_changeEnd #from 02a.winners.losers.Rmd is mean cover change #needed for P/A???

cbind(Species.ord, contribution.value.ordered, as.data.frame(pct_changeEnd)) -> species.dat 
subset(species.dat, select=-c(Species)) -> species.dat
as.numeric(as.character(species.dat$cumsum)) -> species.dat$cumsum

#species.dat is the nmds matrix.
#add a type column based on  species.dat, cumsum > 0.02 which indicates SIMPER
species.dat -> df
df <- df %>% mutate(NewCol = ifelse(cumsum >= 0.02, "simper", "no"))
df2 <- df %>% mutate(simperType = ifelse(NewCol == "simper" & Diff >= 0, "W", "N"))

df2$simperType[75] <- "L"
df2$simperType[29] <- "L"
df2$simperType[2] <- "L"

#pc = data used in nmds: 
pc <- test.traits[-1]
pc2 <- left_join(test.traits, df2, by="species")
m_com = as.matrix(pc)

ano = anosim(m_com, pc2$simperType, distance = "gower", permutations = 9999)

d <- dist(test.traits[-1], method="euclidean") #dissimilarity matrix
clust <- hclust(d, method = "complete", members = NULL)
plot(Ord) #plot ordination points. red points = traits
ordicluster(Ord, clust, prune=3, col = cutree(clust, h=6)) #add vectors 4 colors
clust$labels %in% sim$Species #returns logical use to subset $labels and $heights
clust[clust$labels %in% sim$Species, ]
subset(clust$labels, clust$labels %in% sim$Species)
```

```{r}
species.in.space <- ggplot(species.dat, aes(NMDS1, NMDS2, size=cumsum, color=PctChange, label=species)) +
   geom_point() + 
   scale_color_gradient2(low = "blue", mid = "white",
  high = "red", midpoint = 0) +
  labs(color = "cover change", size= "simper") + 
  theme_bw()
 
species.in.space2 <- species.in.space  +
  geom_text_repel(data=subset(species.dat, cumsum > 0.02), size=2.5, vjust=2, color="black", min.segment.length = unit(0.01, 'lines'),nudge_y = 0) 
```

#Calculate phylogenetic signal of contribution.value (SIMPER)
#Calculate mean pairwise distance between simper species with highest cover change with and without outliers from random draws of same number taxa:
```{r}
species.distMat <- vegdist(test.traits[-1], method="bray") 
listofT <- subset(species.dat$species, species.dat$change > 100)
library(usedist)
subDist <- dist_subset(species.distMat, listofT)

mean(subDist) -> obsDist #0.285566 #
#length(listofT) #15
#ranT <- sample(species.dat$species, 15)
#ranDist <- dist_subset(species.distMat, ranT)
meanRanDists <- numeric(10)
ranDistMat <- vector('list', 10)
namesList <- replicate(10, sample(species.dat$species, 15))

for (i in 1:dim(namesList)[2]) {
   tempDist <- dist_subset(species.distMat, namesList[,i])
   ranDistMat[[i]] <- tempDist
   meanRanDists[i]<- mean(ranDistMat[[i]])
}
hist(meanRanDists)
abline(v=obsDist)
```

```{r}
#phy signal of contribution.value
contribution.value[1:10, 1] -> simperSPP

ReorderData(tree, contribution.value, taxa.names=1) -> dayt
sig <- phylosig(tree, as.numeric(dayt$cumsum), method="lambda", test=TRUE)
sigK <- phylosig(tree2, contable.ordered$test_stat_plug, method="K", test=TRUE)
sig$lambda
prop.sigK$K

#pc2$simperType
dat <- as.data.frame(cbind(pc2$species, pc2$simperType))
dat2 <- ReorderData(tree, dat, taxa.names=1) 

dat2copy <- dat2
dat2copy$V2 <- as.character(dat2copy$V2)
dat2copy$V2[dat2copy$V2 == "W"] <- "1"
dat2copy$V2[dat2copy$V2 == "L"] <- "0"
dat2copy$V2[dat2copy$V2 == "N"] <- "0"

rownames(dat2) <- dat2$V1
dat3<- subset(dat2copy, select=V2)
sig <- phylosig(tree, as.numeric(dat3$V2), method="lambda", test=TRUE)

mytree<-multi2di(tree)
result <- fitDiscrete(mytree, char1 , treeTransform=c("lambda"))
#phylo.d(dat2, tree, names.col = V1)
```

##EXTRA FIGURE
#simper species cover over time, colored by traits:
```{r}
df.coverlong <- melt(df.cover.na, id.vars = c("plot", "year", "type", "phyD", "fd", "Code", "new.rep"), variable.name = "Species")
df.coverlong  <- filter(df.coverlong, value !="NA")
split(df.coverlong, df.coverlong$type) -> df.coverlong2

df.mean.cover <- df.coverlong2$Plug %>%
  group_by(Species, year, type) %>%
  summarise(Mean=mean(value), Sd= sd(value))

dfMcov.traits <-  merge(df.mean.cover, traits.mat[, c("Species", "SDMC", "SLA", "vegetativeHeight", "LDMC", "LNC", "LCC", "LDMC")], by="Species")
dfMcov.traits$logMean <- log(dfMcov.traits$Mean) #Log tranform cover
#species.dat #from simper.figures.Rmd  
simper.df <- species.dat[order(-species.dat$cumsum),]
simperSpecies <- simper.df[1:10,]

simperSpecies.traits <-  left_join(simperSpecies, dfMcov.traits, by.x="species", by.y="Species")

simper.plot <- ggplot(simperSpecies.traits, aes(x=year)) + 
  geom_line(aes(y=logMean, col=vegetativeHeight, group=Species), size=2) + 
  labs(title="Mean Cover of simper species", 
       subtitle="Colored by Trait", 
       caption="", 
       y="(log)Mean Cover by Species", 
       color=NULL) +
  geom_dl(aes(label= Species, y=logMean), method=list(dl.combine("first.points", "last.points"), cex = 0.5))
```
